---
title: "p8105_hw1_msk2265.Rmd"
author: "Mirah"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
# Problem 1

First, I am loading the library and dataset as specified in the homework instructions.
```{r} 
library(moderndive) #load library
data("early_january_weather") #loading data 

```

The following is a description of the dataset that I produced using inline R code.

  * The variables in this dataset include `r colnames(early_january_weather)`. This data is all from the month `r mean(pull(early_january_weather, month))`(January) and the year `r mean(pull(early_january_weather, year))`. The data only consists of `r min(pull(early_january_weather, time_hour))` through `r max(pull(early_january_weather, time_hour))`. 
  
  * The dataset has `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns. 
  
  * The mean temperature is `r mean(pull(early_january_weather, temp))`.

Next, I will make a scatterplot of temperature versus time_hours using the early_January_weather dataset 
```{r}
#Make a scatterplot of temp (y) vs time_hour (x); color points using the humid variable (adding color = ... inside of aes in your ggplot code should help). Describe patterns that are apparent in this plot.

ggplot(early_january_weather, aes(x = temp, y = time_hour, color = humid)) + geom_point()

#export scatterplot to project directory using ggsave
ggsave("Scatterplot tempvtime earlyjanwalther.pdf", height = 4, width = 6)

```

Based on the scatter plot above, it looks like the temperature fluctuates daily. The highest temperature seems to be in the middle of the day and the lowest temperature at the beginning and end of the day. Overall, temperature increases as time went on, meaning the temperature at the beginning of January is lower than the temperature on the 14th, although it dropped again on the 15th. the humidity also increases with time, and it looks like the most human time of the first half of January was between day 10 and day 15. 


# Problem 2 

Instructions: This problem is intended to emphasize variable types and introduce coercion; some awareness of how R treats numeric, character, and factor variables is necessary for working with these data types in practice.

Create a data frame comprised of:

a random sample of size 10 from a standard Normal distribution
a logical vector indicating whether elements of the sample are greater than 0
a character vector of length 10
a factor vector of length 10, with 3 different factor “levels”

```{r}
set.seed(1234) #make sure that the random data stays the same
df = tibble(
  vec_random = rnorm(10), #random sample from standard normal distribution of size 10
vec_logic = vec_random > 0, #logic vector that indicates which of vec_random are grater than 0
vec_character = c("dog", "cat", "mouse", "bunny", "bird", "pig", "cow", "fish", "sheep", "hen"), #character vector i did animals, there werent instructions on what to make this 
vec_factor = factor(c("one", "one", "one", "two", "two", "two", "three", "three","three", "three")), #factor vector with 3 levels
)
```

instructions: Try to take the mean of each variable in your dataframe. What works and what doesn’t?
```{r}
mean(pull(df,vec_random)) #calculating the mean of the vec_random vector in the df dataframe
mean(pull(df,vec_logic)) #calculating the mean of the vec_logic vector in the df dataframe
```

The vector of 10 numbers returns a mean and the logic vector also returns a mean.
``` {r}
mean(pull(df,vec_character))#attempting to calculate the mean of the vec_character vector in the df dataframe
mean(pull(df,vec_factor))#attempting to calculate the mean of the vec_factor vector in the df dataframe
```
The character vector and factor vector do not return a mean. The warning states that the argument is only valid for numerical and logical vectors.


Instructions: In some cases, you can explicitly convert variables from one type to another. Write a code chunk that applies the as.numeric function to the logical, character, and factor variables (please show this chunk but not the output). What happens, and why? Does this help explain what happens when you try to take the mean?
```{r, eval = FALSE}
#showing chunch but not output
as.numeric(pull(df, vec_logic)) #converting vec_logic to numeric
as.numeric(pull(df, vec_character))#converting vec_character to numeric
as.numeric(pull(df, vec_factor))#converting vec_factor to numeric
```
When converting the logical vector, character vector, and factor vector to numeric vectors, the logical and factor vectors create sets of numbers, but the character vector just returns NA. This explains why there is no mean for character variables. For the factor vector, it looks like there is a numerical output but it is numbered in alphabetical order, not by the numbers I actually wrote. When I actually calculate the mean of my factor vector after converting it to numeric, there is a mean.

